---
title: "p8105_hw2_bf2506"
output: github_document
date: "2022-09-29"
---
```{r load package, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
```


### Problem 2

##### 1.Read and clean Mrtrash_data
```{r read Mrtrash_wheel_data, message=FALSE, warning=FALSE}
raw_Mrtrash_wheel_data =  
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N533",
  na = c("", "Decemeber Total", "November Total", "October Total", "September Total", "August Total", "July Total", "June Total", "May Total", "April Total", "March Total", "February Total", "January Total")
)
Mrtrash_wheel_data = drop_na(raw_Mrtrash_wheel_data, Dumpster)

```

```{r rename and integer sportsballs in Mr, message=FALSE, warning=FALSE}

Mrtrash_wheel_data = janitor::clean_names(Mrtrash_wheel_data) %>% rename(volume_cubicyards = volume_cubic_yards) %>% mutate(sports_balls=as.integer(sports_balls)) 

```

##### 2.Read and clean Professortrash_data
```{r read Professortrash_wheel_data, message=FALSE, warning=FALSE}
raw_Professortrash_wheel_data = 
  readxl::read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Professor Trash Wheel", 
  range = "A2:N115", 
  na = c("", "Decemeber Total", "November Total", "October Total", "September Total", "August Total", "July Total", "June Total", "May Total", "April Total", "March Total", "February Total", "January Total")
)

Professortrash_wheel_data = drop_na(raw_Professortrash_wheel_data, Dumpster)

```

```{r rename and integer sportsballs in PRO, message=FALSE, warning=FALSE}

Professortrash_wheel_data = janitor::clean_names(Professortrash_wheel_data) %>% rename(volume_cubicyards = volume_cubic_yards) %>% mutate(sports_balls = as.integer(sports_balls)) 

```


##### 3.Combine two tables
```{r combine tables, message=FALSE, warning=FALSE}
Mrtrash_wheel_data = 
  mutate(Mrtrash_wheel_data, trash_wheel = "Mr_Trash_Wheel")

Professortrash_wheel_data =
  mutate(Professortrash_wheel_data, trash_wheel = "Professor_Trash_Wheel")

Trash_Wheel_data = 
  bind_rows(Mrtrash_wheel_data, Professortrash_wheel_data) %>% relocate(trash_wheel)

```

**Describe the Data form Trash_Wheel_data**
 
The size of the dataset is **`r ncol(Trash_Wheel_data)` variables** and **`r nrow(Trash_Wheel_data)` observations**, showing the summary of trash collection from two trash wheel(Mr.trash wheel and Professor trash wheel). And the total weight of trash collected by Professor Trash Wheel is`r sum(Professortrash_wheel_data$weight_tons)` and the total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(filter(Mrtrash_wheel_data, year == 2020)$sports_balls)`

**15** variables are as following:  
**·trash_wheel**: whether from "Mr_Trash_Wheel" or "	
Professor_Trash_Wheel" <br>
**·dumpster**: the number of dumpsters  <br>
**·month**: the month that collect the data, from Jan to Dec <br>
**·year**: the year that collect the data <br>
**·date**: the date that collect the data <br>
**·weight_tons**: weight in tons <br>
**·volume_cubicyards**: volume in cubic yards <br>
**·plastic_bottles**:  the number of plastic bottles <br>
**·polystyrene**: the number of polystyrene <br>
**·cigarette_butts**: the number of cigarette butts <br> 
**·glass_bottles**: the number of glass bottles <br>
**·grocery_bags**: the number of grocery bags <br>
**·chip_bags**: the number of chip bags <br>
**·sports_balls**：the number of sports balls <br>
**·homes_powered**: the number of homes powered by  electricity generated from these trash <br>
 
### Problem 3

##### 1.Clean the data in pols-month.csv
```{r read and clean  pols_month_data, message=FALSE, warning=FALSE}
pols_month_data =
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "\\-") %>% 
  mutate( year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate( month = month.abb[month]) %>%
  mutate(president = ifelse(prez_gop - prez_dem > 0, "gop", "dem")) %>% 
  select(-c(prez_gop, prez_dem, day))

```

##### 2.Clean the data in snp.csv
```{r read and clean  snp_data, message=FALSE, warning=FALSE}

snp_data =
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE ) %>% 
  select(-day) %>% 
  mutate( year = ifelse( year < 2050, as.integer(year),  as.integer(year - 100) )) %>%  
  arrange(year, month) %>% 
  mutate(month = month.abb[month]) %>% 
  select(year, month, everything())

```

##### 3.Tidy unemployment data

```{r read and clean and tidy unemployment_data, message=FALSE, warning=FALSE}

unemployment_data = 
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "unemployment_percentage") %>% 
  mutate(month = sub("(.)", "\\U\\1", month, perl = TRUE), year = as.integer(year))

```

##### 4.Joint pols_month/snp/unemployment tables

```{r joint tables, message=FALSE, warning=FALSE}

pols_snp_unemployment_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month")) %>% 
  left_join(unemployment_data, by = c("year", "month")) %>% 
  select(year, month, close, unemployment_percentage, president, everything())

```
##### 5.Describe these tables

**pols_month_data**  is related to the number of national politicians who are democratic or republican per month from January 1947 to June 2015 and contains 822 observations of 9 variables ("year", "month，"president", "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem")
 
**snp_data** is related to Standard & Poor’s stock market index (a representative measure of the stock market as a whole) per month from January 1950 to July 2015 and contains 787 observations of 3 variables ("year", "month", "close")
       
**unemployment_data** is related to percentage of unemployment per month from January 1948 to December 2015 and contains 816 observations of 3 variables ("year", "month", "unemployment_percentage")

**pols_snp_unemployment_data** is merged by pols_month_data, snp_data and unemployment_data, related to  the number of national politicians who are democratic or republican/S&P index/
unemployment percentage. The 3 in 1 table contains 822 observations and 11 variables per year from January 1947 to June 2015. Some key variables are as following:
 
**year** : the year of the observations on that row <br>
**month** : the month of the observations on that row <br>
**close** : the closing values of the S&P stock index <br>
**unemployment** : the percentage of unemployment <br>
 **president** : whether the president was dem or gop <br>


### Problem 1

##### 1.Read and clean and select NYC_subway_data
```{r read and clean and select NYC_subway_data}
NYC_subway_data = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

##### 2.Describe NYC_subway_data
I import data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.(`r nrow(NYC_subway_data)`rows x `r ncol(NYC_subway_data)`columns)

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 
 
##### 3.Calculate number of distinct stations
```{r number of distinct stations, message=FALSE, warning=FALSE}
NYC_subway_data %>% 
  select(station_name, line) %>% 
  distinct
```

##### 4.Calculate number of ADA stations
```{r number of ADA stations, message=FALSE, warning=FALSE}
NYC_subway_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

##### 5.Proportion of entr/exi without vending allow entrance
```{r message=FALSE, warning=FALSE}
NYC_subway_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

##### 6.Stations serving A train/ADA
```{r message=FALSE, warning=FALSE}
NYC_subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYC_subway_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```